



This script creates the following .rds files:
utilities, plants, generators, cooling, net_generation, and cooling_detail

This script will read in data from EIA forms 860 and 923
To run it, you need to download the forms and unzip them.

```{r}
library(readxl)
library(tidyverse)
library(lubridate)
library(zoo)
```

## Electric Utilities

```{r}
############# FORM 860
## 2001 - 2003 : in .dbf format

## 2004
## UtilY04 have PlantY04 which have GenY04
## MFExist04 has info regarding multi-fuel plants
## PRGenY04 has info regarding proposed plants
## MFProp04 has info regarding proposed multi-fuel plants
## PCGenY04 has info regarding proposed changes to existing plants
## OwnerY04 has info regarding generators that are owned by someone other than the operator.
## Layout describes the other workbooks.
## Each workbook aside from Layout is a single worksheet.

## 2005 - 2006 : same files as 2004, plus Intcon05, which has data on new interconnection costs.

## 2007 - 2008 : doesn't have PCGen
## does have F860_Schedule6 with a bunch of sheets on the boilers
## and another book to explain the F860_Schedule6 sheets.

## 2009 - 2012 : different names from previous.
## apparently broke up the Schedule6 sheets into two Enviro sheets.

## 2013 - 2017 : different file names from previous.
## generator file is multi-sheet. 'Operable', 'Proposed', 'Retired and Cancelled'
## I only need 'Operable'
## includes wind and solar workbooks.

## TODO: for the most recent year, look at the worksheet named 'Proposed'.

readxl::read_excel("downloaded excel spreadsheets/form 860/eia8602013/1___Utility_Y2013.xlsx")
```

```{r}
## This for look will read in tables (utlity, plant, and generator) for each year
for(i in 2013:2023) {
  folder <- paste0('downloaded excel spreadsheets/form 860/eia860', i, '/')

  utility <- readxl::read_xlsx(paste0(folder, '1___Utility_Y', i,'.xlsx'),
                       sheet=1, skip=1)
  utility$Year <- i
  assign(paste0('utility', i), utility)

  plant <- readxl::read_xlsx(paste0(folder, '2___Plant_Y', i, '.xlsx'),
                     sheet=1, skip=1, guess_max=5000)
  plant$Year <- i
  assign(paste0('plant', i), plant)

  generator <- readxl::read_xlsx(paste0(folder, '3_1_Generator_Y', i, '.xlsx'),
                         sheet=1, skip=1, guess_max=10000)
  generator$Year <- i
  assign(paste0('generator', i), generator)

  rm(utility, plant, generator, folder, i)
}
```

```{r}
names(utility2013) <- names(utility2016)
names(utility2014) <- names(utility2016)
names(utility2015) <- names(utility2016)
utility2017 <- select(utility2017, -Caution)

utility2023 <- dplyr::mutate(utility2023, Zip = as.character(Zip))
```

```{r}
utilities <- bind_rows(
  utility2013, utility2014, utility2015, utility2016, utility2017, utility2018,
  utility2019, utility2020, utility2021, utility2022, utility2023)

usethis::use_data(utilities, overwrite=T)
```

```{r}
rm(utility2013, utility2014, utility2015, utility2016, utility2017, utility2018,
   utility2019, utility2020,utility2021,utility2022, utility2023)
```


## Power Plants

```{r}
plant2017 <- select(plant2017, -Caution)
plants <- bind_rows(
  plant2013, plant2014, plant2015, plant2016,
  plant2017, plant2018, plant2019, plant2020, plant2021, plant2022, plant2023)
```

```{r}
rm(plant2013, plant2014, plant2015, plant2016,
   plant2017, plant2018, plant2019, plant2020, plant2021, plant2022, plant2023)

usethis::use_data(plants, overwrite=T)
```

## Electricity Generators
```{r}
generator2017 <- select(generator2017, -Caution)

generators0 <- bind_rows(
  generator2013, generator2014, generator2015, generator2016, generator2017, 
  generator2018, generator2019, generator2020, generator2021, generator2022, 
  generator2023)
```

```{r}
rm(generator2013, generator2014, generator2015, generator2016, generator2017, 
   generator2018, generator2019, generator2020, generator2021, generator2022,
   generator2023)
```

Consolidate and decode some of the columns in the Generators table
Decode prime movers
Decode energy sources
Reduce Date Columns
Reduce fuel columns

```{r}
PrimeMovers <- readxl::read_xlsx('Code Sheets_2017.xlsx', sheet=2)
EnergySources <- readxl::read_xlsx('Code Sheets_2017.xlsx', sheet=1) %>%
  select(`Energy Source Code`, `Energy Source Description`)
```

```{r}
generators <- generators0 %>%
  EIAdata::join_and_drop(., PrimeMovers, by='Prime Mover', rename='Prime Mover') %>%
  EIAdata::join_and_drop(., EnergySources, by=c('Energy Source 1'='Energy Source Code'),
                rename=c("PrimaryFuelType", "PrimaryFuel")) %>%
  EIAdata::join_and_drop(EnergySources, c('Energy Source 2'='Energy Source Code'),'Fuel2') %>%
  EIAdata::join_and_drop(EnergySources, c('Energy Source 3'='Energy Source Code'),'Fuel3') %>%
  EIAdata::join_and_drop(EnergySources, c('Energy Source 4'='Energy Source Code'),'Fuel4') %>%
  EIAdata::join_and_drop(EnergySources, c('Energy Source 5'='Energy Source Code'),'Fuel5') %>%
  EIAdata::join_and_drop(EnergySources, c('Energy Source 6'='Energy Source Code'),'Fuel6') %>%
  EIAdata::join_and_drop(EnergySources, c('Startup Source 1'='Energy Source Code'),'Start1') %>%
  EIAdata::join_and_drop(EnergySources, c('Startup Source 2'='Energy Source Code'),'Start2') %>%
  EIAdata::join_and_drop(EnergySources, c('Startup Source 3'='Energy Source Code'),'Start3') %>%
  EIAdata::join_and_drop(EnergySources, c('Startup Source 4'='Energy Source Code'),'Start4') %>%
  EIAdata::monthyear_to_date('Month Uprate or Derate Completed', 'Year Uprate or Derate Completed',
                    'Date Uprate or Derate Completed') %>%
  EIAdata::monthyear_to_date('Operating Month', 'Operating Year', 'Operating Date') %>%
  EIAdata::monthyear_to_date('Planned Retirement Month', 'Planned Retirement Year', 'Planned Retirement Date') %>%
  EIAdata::monthyear_to_date('Planned Uprate Month', 'Planned Uprate Year', 'Planned Uprate Date') %>%
  EIAdata::monthyear_to_date('Planned Derate Month', 'Planned Derate Year', 'Planned Derate Date') %>%
  EIAdata::monthyear_to_date('Planned Repower Month', 'Planned Repower Year', 'Planned Repower Date') %>%
  EIAdata::monthyear_to_date('Other Modifications Month', 'Other Modifications Year', 'Planned Other Modifications Date') %>%
  EIAdata::paste_columns(c('Start1', 'Start2', 'Start3', 'Start4'), 'StartupFuels') %>%
  EIAdata::paste_columns(c('Fuel2', 'Fuel3', 'Fuel4', 'Fuel5', 'Fuel6'), 'SecondaryFuels')
```

```{r}

usethis::use_data(generators, overwrite=T)

# for(i in 2013:2016) {
#   folder <- paste0('form 860\\eia860', i, '\\')
#
#   EnviroAssoc <- read_xlsx(paste0(folder, '6_1_EnviroAssoc_Y', i,'.xlsx'),
#                        sheet=1, skip=1)
#   EnviroAssoc$Year <- i
#   assign(paste0('EnviroAssoc', i), EnviroAssoc)
#
#   EnviroEquip <- read_xlsx(paste0(folder, '6_2_EnviroEquip_Y', i, '.xlsx'),
#                      sheet=1, skip=1)
#   EnviroEquip$Year <- i
#   assign(paste0('EnviroEquip', i), EnviroEquip)
#
#   rm(EnviroAssoc, EnviroEquip, folder, i)
# }
#
# EnviroAssoc2013_2016 <- bind_rows(
#   EnviroAssoc2013, EnviroAssoc2014, EnviroAssoc2015, EnviroAssoc2016)
# rm(EnviroAssoc2013, EnviroAssoc2014, EnviroAssoc2015, EnviroAssoc2016)
#
# EnviroEquip2013_2016 <- bind_rows(
#   EnviroEquip2013, EnviroEquip2014, EnviroEquip2015, EnviroEquip2016)
# rm(EnviroEquip2013, EnviroEquip2014, EnviroEquip2015, EnviroEquip2016)

```

```{r}
######### form 923.
## It has a lot of pages/'schedules'/sheets in the yearly workbooks
## I compiled the two sheets I thought were relevant from each year into two workbooks.
## (manually). So, could be good to write a function to read in all the sheets...

## havent done this for 2023 data yet. got an error when I run this chunk.

for(i in  2012:2022) {
  netGen <- read_xlsx(
    'downloaded excel spreadsheets\\f923_Page4_Generators.xlsx',
    sheet=as.character(i), skip=5, guess_max=10000)
  names(netGen) <- c(
    'PlantID', 'Combined Heat and Power', 'Plant', 'Operator', 'OperatorID',
    'State', 'CensusRegion', 'NERCRegion', 'NAICS', 'SectorNumber', 'Sector',
    'GeneratorID', 'PrimeMover', 'Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun',
    'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'YearToDate', 'Year')
  assign(paste0('netGen', i), netGen)

  cooling <- read_xlsx(
    'downloaded excel spreadsheets\\f923_Schedule8_Cooling.xlsx',
    sheet=as.character(i), skip=4, guess_max=10000)
  names(cooling) <- c(
    'Year', 'Month', 'PlantID', 'CoolingID', 'CoolingType', 'CoolingStatus', 'HoursInService',
    'Chlorine (thousand lbs)', 'Diversion (gpm)', 'Withdrawal (gpm)', 'Discharge (gpm)',
    'Consumption (gpm)', 'FlowMethod', 'IntakeAvgTemp (F)', 'IntakeMaxTemp', 'DischargeAvgTemp',
    'DischargeMaxTemp', 'TempMethod', 'Diversion (MGal)', 'Withdrawal (MGal)', 'Discharge (MGal)',
    'Consumption (MGal)')
  assign(paste0('cooling', i), cooling)

  rm(netGen, cooling, i)
}
```

```{r}
## f923_Schedule8_Cooling
## I'm not sure I need this, use cooling_detail instead ?
cooling2019 <- dplyr::mutate(cooling2019, HoursInService = as.character(HoursInService))
cooling <- bind_rows(
  cooling2012, cooling2013, cooling2014, cooling2015,
  cooling2016, cooling2017, cooling2018, cooling2019, cooling2020, cooling2021, cooling2022)
rm(cooling2012, cooling2013, cooling2014, cooling2015,
   cooling2016, cooling2017, cooling2018, cooling2019, cooling2020, cooling2021, cooling2022)

cooling <- cooling %>%
  EIAdata::monthyear_to_date(.) %>%
  dplyr::mutate(dplyr::across(
    c("HoursInService", "IntakeAvgTemp (F)", "IntakeMaxTemp",
      "DischargeAvgTemp", "DischargeMaxTemp", "Chlorine (thousand lbs)",
      "Diversion (MGal)", "Withdrawal (MGal)", "Discharge (MGal)", "Consumption (MGal)"),
    function (x) as.numeric(dplyr::na_if(x, '.'))))

usethis::use_data(cooling, overwrite=T)
```

# Net Generation
From Form 923, Page 4, "Generators."

```{r}
### f923_Page4_Generators
years <- c("2020", "2021", "2022")

# Create an empty list to store the data frames
netGen_list <- list()

# Loop through each year
for (year in years) {
  # Read the Excel file
  netGen <- read_xlsx(
    'data-raw\\downloaded excel spreadsheets\\f923_Page4_Generators.xlsx',
    sheet = year, skip = 5, guess_max = 10000)

  # Assign column names
  names(netGen) <- c(
    'PlantID', 'Combined Heat and Power', 'Plant', 'Operator', 'OperatorID',
    'State', 'CensusRegion', 'NERCRegion', 'NAICS', 'SectorNumber', 'Sector',
    'GeneratorID', 'PrimeMover', 'Respondent Frequency', 'Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun',
    'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'YearToDate', 'BalancingAuthorityCode', 'Year')

  # Store the data frame in the list
  netGen_list[[year]] <- netGen
}

# Access the data frames using names
netGen2020 <- netGen_list[["2020"]]
netGen2021 <- netGen_list[["2021"]]
netGen2022 <- netGen_list[["2022"]]

# netGen2020 <- read_xlsx(
#   'data-raw\\downloaded excel spreadsheets\\f923_Page4_Generators.xlsx',
#   sheet="2020", skip=5, guess_max=10000)
#
# names(netGen2020) <- c(
#   'PlantID', 'Combined Heat and Power', 'Plant', 'Operator', 'OperatorID',
#   'State', 'CensusRegion', 'NERCRegion', 'NAICS', 'SectorNumber', 'Sector',
#   'GeneratorID', 'PrimeMover', 'Respondent Frequency', 'Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun',
#   'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'YearToDate', 'BalancingAuthorityCode', 'Year')
```

```{r}
net_generation <- bind_rows(
  netGen2012, netGen2013, netGen2014, netGen2015,
  netGen2016, netGen2017, netGen2018, netGen2019, netGen2020, netGen2021, netGen2022)

net_generation <- net_generation %>%
  EIAdata::join_and_drop(
    ., PrimeMovers, by=c('PrimeMover'='Prime Mover'), rename='Prime Mover')
```

```{r}
usethis::use_data(net_generation, overwrite=T)

rm(netGen2012, netGen2013, netGen2014, netGen2015,
   netGen2016, netGen2017, netGen2018, netGen2019, netGen2020, netGen2021, netGen2022)
```

# Cooling Detail
```{r}

## I downloaded these sheets from https://www.eia.gov/electricity/data/water/
## and compiled them in to a single workbook (it is too big, probably should have kept them separate)
## and deleted some whitespace in the title of the first column heading ('Utility ID')
## and deleted two blank rows at the top.

cj14 <- readxl::read_xlsx(
  'data-raw\\downloaded excel spreadsheets\\cooling detail\\cooling_detail_2014.xlsx',
  guess_max=10000) %>%
  dplyr::mutate(`Steam Plant Type`=as.numeric(`Steam Plant Type`))

cj15 <- readxl::read_xlsx(
  'data-raw\\downloaded excel spreadsheets\\cooling detail\\cooling_detail_2015.xlsx',
  guess_max=10000) %>%
  dplyr::mutate(
    `Other Gas Consumption (MMBTU)` = as.numeric(`Other Gas Consumption (MMBTU)`),
    `Steam Plant Type`=as.numeric(`Steam Plant Type`))

cj16 <- readxl::read_xlsx(
  'data-raw\\downloaded excel spreadsheets\\cooling detail\\cooling_detail_2016.xlsx',
  skip=2, guess_max=10000) %>%
  dplyr::mutate(
    `Other Gas Consumption (MMBTU)` = as.numeric(`Other Gas Consumption (MMBTU)`),
    `Steam Plant Type`=as.numeric(`Steam Plant Type`))

cj17 <- readxl::read_xlsx(
  'data-raw\\downloaded excel spreadsheets\\cooling detail\\cooling_detail_2017.xlsx',
  skip=2, guess_max=10000) %>%
  dplyr::mutate(Year=as.numeric(Year), Month=as.numeric(Month))

cj18 <- readxl::read_xlsx(
  'data-raw\\downloaded excel spreadsheets\\cooling detail\\cooling_detail_2018.xlsx',
  skip=2, guess_max=10000) %>%
  dplyr::mutate(Year=as.numeric(Year), Month=as.numeric(Month))

cj19 <- readxl::read_xlsx(
  'data-raw\\downloaded excel spreadsheets\\cooling detail\\cooling_detail_2019.xlsx',
  skip=2, guess_max=10000) %>%
  dplyr::mutate(Year=as.numeric(Year), Month=as.numeric(Month))

cj20 <- readxl::read_xlsx(
  'data-raw\\downloaded excel spreadsheets\\cooling detail\\cooling_detail_2020.xlsx',
  skip=2, guess_max=10000) %>%
  dplyr::mutate(Year=as.numeric(Year), Month=as.numeric(Month))

cj21 <- readxl::read_xlsx(
  'data-raw\\downloaded excel spreadsheets\\cooling detail\\cooling_detail_2021.xlsx',
  skip=2, guess_max=10000) %>%
  dplyr::mutate(Year=as.numeric(Year), Month=as.numeric(Month))

cj22 <- readxl::read_xlsx(
  'data-raw\\downloaded excel spreadsheets\\cooling detail\\cooling_detail_2022.xlsx',
  skip=2, guess_max=10000) %>%
  dplyr::mutate(Year=as.numeric(Year), Month=as.numeric(Month))

```

```{r}
cooling_detail0 <-
  list(cj22,cj21,cj20, cj19, cj18, cj17, cj16, cj15, cj14) %>%
  lapply(
    FUN= function(x) {
      names(x)[1]<-"Utility ID"
      dplyr::select(x, -`...59`, -`...60`)}) %>%
  bind_rows()

cooling_detail <- cooling_detail0 %>%
  EIAdata::monthyear_to_date('Cooling Inservice Month', 'Cooling Inservice Year', 'Cooling Inservice Date') %>%
  EIAdata::monthyear_to_date('Boiler Retirement Month', 'Boiler Retirement Year', 'Boiler Retirement Date') %>%
  EIAdata::monthyear_to_date('Boiler Inservice Month', 'Boiler Inservice Year', 'Boiler Inservice Date') %>%
  EIAdata::monthyear_to_date('Generator Retirement Month', 'Generator Retirement Year', 'Generator Retirement Date') %>%
  EIAdata::monthyear_to_date('Generator Inservice Month', 'Generator Inservice Year', 'Generator Inservice Date') %>%
  EIAdata::join_and_drop(PrimeMovers, c('Generator Prime Mover Code'='Prime Mover'), 'Prime Mover') %>%
  EIAdata::join_and_drop(EnergySources, c('Generator Primary Energy Source Code'='Energy Source Code'),
                c("PrimaryFuelType", "PrimaryFuel")) %>%
  left_join(select(plants, 1:3, Latitude, Longitude, Year)) %>%
  left_join(select(generators, `Plant Code`, `Generator ID`, Year,
                   `Nameplate Capacity (MW)`, `Nameplate Power Factor`,
                   `Summer Capacity (MW)`, `Winter Capacity (MW)`,
                   `Minimum Load (MW)`)) %>%
  EIAdata::monthyear_to_date(.)

```

```{r}
usethis::use_data(cooling_detail, overwrite=T)

```

```{r}

```



